\pkg{NeuralNetTools}: Visualization and Analysis Tools for Neural Networks

### 2018-05-04 final_JSS_edits/paper.tex
messing with reproducible stuff from jss

First, the appropriate month and airline carrier are selected, all
explanatory variables are scaled and centered, and the response variable is
scaled to 0--1.
%
\begin{Schunk}
\begin{Sinput}
R> library("nycflights13")
R> library("dplyr")
R> tomod <- filter(flights, month == 12 & carrier == "UA") %>% 
+    select(arr_delay, dep_delay, dep_time, arr_time, air_time,
+      distance) %>% mutate_each(funs(scale), -arr_delay) %>% 
+    mutate_each(funs(as.numeric), -arr_delay) %>% 
+    mutate(arr_delay = scales::rescale(arr_delay, to = c(0, 1))) %>% 
+    data.frame
\end{Sinput}
\end{Schunk}
%
Then, a standard MLP with five hidden nodes was created with the
\pkg{nnet} package to model the effects of selected variables on
arrival delays. The entire dataset is used for the example but
separate training and validation datasets should be used in practice.
%
\begin{Schunk}
\begin{Sinput}
R> library("nnet")
R> mod <- nnet(arr_delay ~ ., size = 5, linout = TRUE, data = tomod,
+    trace = FALSE)  
\end{Sinput}
\end{Schunk}
%
The default output is limited to structural information about the model and methods for model predictions (see \code{str(mod)} and \code{?predict.nnet}). Using functions from \pkg{NeuralNetTools}, a more comprehensive understanding of the relationships between the variables is illustrated.
%
\begin{Schunk}
\begin{Sinput}
R> plotnet(mod)
R> garson(mod)
R> olden(mod)
R> lekprofile(mod, group_vals = 5)
R> lekprofile(mod, group_vals = 5, group_show = TRUE)
\end{Sinput}
\end{Schunk}
%
\Cref{fig:flightplots} shows the information about arrival delays that
can be obtained with the functions in \pkg{NeuralNetTools}. The
\ac{nid} (\ref{fig:flightplotsa}) shows the model structure and can be
used as a general characterization of the relationships between
variables.  For example, most of the connection weights from input
node I5 are negative (gray), suggesting that distance traveled has an
opposing relationship with arrival delays. Similarly, a large positive
weight is observed between I3 and H1, suggesting arrival time is
positively associated with arrival delays.  However, interpreting
individual connection weights between layers is challenging.
Figures~\ref{fig:flightplotsb} and \ref{fig:flightplotsc} provide more
quantitative descriptions using information from both the \ac{nid} and
model predictions.  \Cref{fig:flightplotsb} shows variable importance
using the \code{garson} and \code{olden} algorithms.  The
\code{garson} function suggests time between destinations
(\code{air\_time}) has the strongest relationship with arrival delays,
similar to a strong positive association shown with the \code{olden}
method. However, the \code{garson} function shows arrival time
(\code{arr\_time}) as the third most important variable for arrival
delays, whereas this is ranked highest by the \code{olden}
function. Similar discrepancies between the two methods are observed
for other variables, which are explained below. Finally, results from
the \code{lekprofile} function (Figure~\ref{fig:flightplotsc}) confirm those
in Figure~\ref{fig:flightplotsb}, with the addition of non-linear responses
that vary by different groupings of the data.  Values for each
variable in the different unevaluated groups (based on clustering)
show that there were no obvious patterns between groups with the
exception being group four that generally had early arrival times and
long departure delays.

>>>>

First, the appropriate month and airline carrier are selected, all
explanatory variables are scaled and centered, and the response variable is
scaled to 0--1.
%
\begin{Schunk}
\begin{Sinput}
R> library("nycflights13")
R> library("dplyr")
R> tomod <- filter(flights, month == 12 & carrier == "UA") %>% 
+    select(arr_delay, dep_delay, dep_time, arr_time, air_time,
+      distance) %>% mutate_each(funs(scale), -arr_delay) %>% 
+    mutate_each(funs(as.numeric), -arr_delay) %>% 
+    mutate(arr_delay = scales::rescale(arr_delay, to = c(0, 1))) %>% 
+    data.frame
\end{Sinput}
\end{Schunk}
%
Then, a standard MLP with five hidden nodes was created with the
\pkg{nnet} package to model the effects of selected variables on
arrival delays. The entire dataset is used for the example but
separate training and validation datasets should be used in practice.
%
\begin{Schunk}
\begin{Sinput}
R> library("nnet")
R> mod <- nnet(arr_delay ~ ., size = 5, linout = TRUE, data = tomod,
+    trace = FALSE)  
\end{Sinput}
\end{Schunk}
%
The default output is limited to structural information about the model and methods for model predictions (see \code{str(mod)} and \code{?predict.nnet}). Using functions from \pkg{NeuralNetTools}, a more comprehensive understanding of the relationships between the variables is illustrated.
%
\begin{Schunk}
\begin{Sinput}
R> plotnet(mod)
R> garson(mod)
R> olden(mod)
R> lekprofile(mod, group_vals = 5)
R> lekprofile(mod, group_vals = 5, group_show = TRUE)
\end{Sinput}
\end{Schunk}
%
\Cref{fig:flightplots} shows the information about arrival delays that
can be obtained with the functions in \pkg{NeuralNetTools}. The
\ac{nid} (\ref{fig:flightplotsa}) shows the model structure and can be
used as a general characterization of the relationships between
variables.  For example, most of the connection weights from input
nodes I2 and I5 are strongly negative (gray), suggesting that departure 
time and distance traveled has an opposing relationship with arrival 
delays. Similarly, large positive weights are observed for I3 and I4, 
suggesting arrival time and time in the air are positively associated 
with arrival delays.  However, interpreting individual connection weights 
between layers is challenging. Figures~\ref{fig:flightplotsb} and 
\ref{fig:flightplotsc} provide more quantitative descriptions using 
information from both the \ac{nid} and model predictions.  
\Cref{fig:flightplotsb} shows variable importance using the \code{garson} 
and \code{olden} algorithms.  The \code{garson} function suggests time 
between destinations (\code{air\_time}) has the strongest relationship 
with arrival delays, similar to a strong positive association shown 
with the \code{olden} method. However, the \code{garson} function 
shows arrival time (\code{arr\_time}) as the third most important 
variable for arrival delays, whereas this is ranked highest by the \code{olden}
function. Similar discrepancies between the two methods are observed
for other variables, which are explained below. Finally, results from
the \code{lekprofile} function (Figure~\ref{fig:flightplotsc}) confirm those
in Figure~\ref{fig:flightplotsb}, with the addition of non-linear responses
that vary by different groupings of the data.  Values for each
variable in the different unevaluated groups (based on clustering)
show that there were no obvious patterns between groups, with the
exception being group one that generally had longer times in the air 
and greater distance travelled.

### 2018-05-04 nnt_manu.Rnw
messing with reproducible stuff from jss

First, the apropriate month and airline carrier were selected, all explanatory variables were normalized, and the response variable was standardized to 0--1.
<<eval = T>>=
library("nycflights13")
library("dplyr")
tomod <- filter(flights, month == 12 & carrier == 'UA') %>% 
  select(arr_delay, dep_delay, dep_time, arr_time, air_time, distance) %>%
  mutate_each(funs(scale), -arr_delay) %>% 
  mutate_each(funs(as.numeric), -arr_delay) %>% 
  mutate(arr_delay = scales::rescale(arr_delay, to = c(0, 1))) %>% 
  data.frame
@
Then, a standard \ac{mlp} with five hidden nodes was created with the \pkg{nnet} package to model the effects of selected variables on arrival delays. The entire dataset is used for the example but separate training and validation datasets should be used in practice.  
<<eval = F, eval = T, results = hide>>=
library(nnet)
mod <- nnet(arr_delay ~ ., size = 5, linout = TRUE, data = tomod, trace = F)  
@
The default output is limited to structural information about the model and methods for model predictions (see \code{str(mod)} and \code{?predict.nnet}). Using functions from \pkg{NeuralNetTools}, a more comprehensive understanding of the relationships between the variables is illustrated.
<<eval = F>>=
plotnet(mod)
garson(mod)
olden(mod)
lekprofile(mod, group_vals = 5)
lekprofile(mod, group_vals = 5, group_show = TRUE)
@
<<echo = F, results = hide>>=
library(nnet)
mod <- nnet(arr_delay ~ ., size = 5, linout = TRUE, data = tomod, trace = F)

>>>>

First, the apropriate month and airline carrier were selected, all explanatory variables were normalized, and the response variable was standardized to 0--1.
<<eval = T>>=
library("nycflights13")
library("dplyr")
tomod <- filter(flights, month == 12 & carrier == 'UA') %>% 
  select(arr_delay, dep_delay, dep_time, arr_time, air_time, distance) %>%
  mutate_each(funs(scale), -arr_delay) %>% 
  mutate_each(funs(as.numeric), -arr_delay) %>% 
  mutate(arr_delay = scales::rescale(arr_delay, to = c(0, 1))) %>% 
  data.frame
@
Then, a standard \ac{mlp} with five hidden nodes was created with the \pkg{nnet} package to model the effects of selected variables on arrival delays. The entire dataset is used for the example but separate training and validation datasets should be used in practice.  
<<eval = F, results = hide>>=
library(nnet)
mod <- nnet(arr_delay ~ ., size = 5, linout = TRUE, data = tomod, trace = F)  
@
The default output is limited to structural information about the model and methods for model predictions (see \code{str(mod)} and \code{?predict.nnet}). Using functions from \pkg{NeuralNetTools}, a more comprehensive understanding of the relationships between the variables is illustrated.
<<eval = F>>=
plotnet(mod)
garson(mod)
olden(mod)
lekprofile(mod, group_vals = 5)
lekprofile(mod, group_vals = 5, group_show = TRUE)
@
<<echo = F, results = hide>>=
library(nnet)
mod <- nnet(arr_delay ~ ., size = 5, linout = TRUE, data = tomod, trace = F)
